---
title       : The Survival of Passengers on the Titanic
subtitle    : 
author      : Patricia Crain
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow     # 
widgets     : [bootstrap, shiny, interactive] 
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## The Sinking of the Titanic

<img height=200 src=titanic.png>

It was one of the biggest tragedies of the Western world, causing worldwide shock at the loss of life and outrage at the negligence that had contributed to the disaster. On April 15, 1912, three hours after striking an iceberg, the RMS Titanic sank into the frigid cold of the north Atlantic. Of the 2201 passengers aboard the Titanic when it embarked from Southampton, England four days earlier, only 711 (32%) survived. 

--- .class #id 

## Report on the Loss of the Titanic
Data on the loss of life was collected by the British Board of Trade and published in a report on July 30, 1912. The variables recorded were Class, Sex, Age and Survived (yes/no). 

```{r echo=FALSE,message=FALSE, results-asis}
require(datasets)
require(reshape2)
require(knitr)
titanic <- data.frame(Titanic)
kable(head(titanic,9), format="pandoc")
titanic <- melt(titanic[,c(1,4,5)])
titanic <- dcast(titanic, Class~Survived, sum)
```

--- .class #id 

## Using GoogleVis to Visualize the Data
This interactive plot allows the user to explore the survival rates by Class of passenger, revealing a higher proportion of first and second class passengers were saved when compared to the number of third class and crew who were saved. 

```{r results='asis', echo=FALSE, message=FALSE}
require(googleVis)
p <- gvisColumnChart(titanic, options=list(width=800, height=250, isStacked=TRUE, title="Survival of Passengers on the Titanic by Class", hAxes="[{title:'Class of Passenger'}]"))
print(p, tag='chart')
```

--- .class #id 

## More Insight Into the Data
What about the other variables in the data? To acheive better insight into the survival rate of passengers, it would be nice to see the impact of Sex and Age on survival. 

To this end, we created 2 filters which allow the user to select each Sex (male/female) and/or each Age (adult/child) and to see this subset of data in the interactive plot. 

--- .class #id 

## The Final App
The completed app has two dropdowns in a side panel. The main panel contains the interactive plot which changes as the dropdowns are changed.

<img height=200 src=screenshot-app.png>

To use the app click [here](https://patriciacrain.shinyapps.io/titanicapp/).

For more info about the sinking of the Titanic click [here](http://www.titanicinquiry.org/BOTInq/BOTReport/botRep01.php).

